---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)

dec_data <- read_rds("decathlon.rds")

library(janitor)

```

```{r}
row.names(dec_data)

dec_data_clean <- clean_names(dec_data)

dec_data_clean


```

# Who had the longest long jump?
```{r}
arrange(dec_data_clean, desc(long_jump))
```
# What was the average 100m time in each competition?

```{r}
mean(dec_data_clean[["x100m"]])
dec_data_100m_og <- filter(dec_data_clean, competition == "OlympicG")
dec_data_100m_deca <- filter(dec_data_clean, competition == "Decastar")
mean(dec_data_100m_og[["x100m"]])
```
# Olympic games time 

```{r}
mean(dec_data_100m_deca[["x100m"]])
```
# Decastar time

Who had the highest total points across both competitions?

```{r}
arrange(dec_data_clean, desc(points))
```
# What was the shot-put scores for the top three competitors in each competition?

```{r}
arrange(dec_data_100m_og, desc(shot_put))
```


```{r}
arrange(dec_data_100m_deca, desc(shot_put))
```
What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
sub_50 <- filter(dec_data_clean, x400m < 50)

mean(sub_50[["points"]])
```
```{r}
above_50 <- filter(dec_data_clean, x400m > 50)

mean(above_50[["points"]])
```






